---
title: "Cancellation Factors"
output: html_notebook
---

```{r}
  source("clean_data.R")
  bookings <- get_cleaned_bookings()          # Supplementary data set
  reservations <- get_cleaned_reservations()  # Primary data set
  
  reservations$order <- rownames(reservations) # for reconstruction after subsetting
  library(ggplot2)
```


Here, the data is being plotted for cancellations against the date and price. 

```{r}

#plotting data from cancelled bookings and date

ggplot(data = bookings, mapping = aes(x = reservation_status_date, y = adr, fill = is_canceled)) + geom_smooth(method = "lm", formula = y~x) + xlab("Date of Stay") + ylab("Average Room Price") 

```


Subsetting plots for Reservation set. 


```{r}
# subset data
average_price <- mean(reservations$avg_price_per_room)
lower_price_room <- subset.data.frame(reservations,
    select = c("arrival_month", "total_nights", "booking_status",
               "avg_price_per_room", "total_rev"))
lower_price_room <- filter(lower_price_room,
                           lower_price_room$avg_price_per_room <= average_price)


upper_price_room <- subset.data.frame(reservations,
    select = c("arrival_month", "total_nights", "booking_status", 
               "avg_price_per_room", "arrival_date_full", "total_rev")) 
upper_price_room <- filter(upper_price_room, 
                           upper_price_room$avg_price_per_room > average_price)



```


Attempting to find the total of cancellations for each date; however, it was not possible. Thus, the data is now reflecting months. 



```{r}
#Find total cancellations per day for each upper and lower prices 

lower_price_room_date <- arrange(lower_price_room, desc(arrival_month))
total_cancellations_lower <- rep(1, times = 12)
for(n in 1:12){
 lower_price_room_filt <- filter(lower_price_room_date, arrival_month == n)
 for (m in 1:length(lower_price_room_filt$arrival_month)){
   if(lower_price_room_filt$booking_status[m] == 1){
     total_cancellations_lower[n] <- total_cancellations_lower[n] + 1
   }
 }
}
lower_price_room_new <- subset(lower_price_room, )



```



Plotting Data for Reservations dataset in stacked bar charts. 


```{r}

ggplot(data = lower_price_room,
       mapping = aes(x = arrival_month,
                     y = total_nights,
                     fill = booking_status)) +
  geom_bar(position = "fill", stat = "identity") +
  xlab("Month of Stay") +
  ylab("Total Nights") +
  ggtitle(label = "Lower Price Room's Cancellations") +
  theme_bw() +
  guides(fill = guide_legend(title = "Reservation Status"))


ggplot(data = upper_price_room,
       mapping = aes(x = arrival_month,
                     y = total_nights,
                     fill = booking_status)) +
  geom_bar(position = "fill", stat = "identity") +
  xlab("Date of Stay") +
  ylab("Total Nights") +
  ggtitle(label = "Upper Price Room's Cancellations") +
  theme_bw() +
  guides(fill = guide_legend(title = "Reservation Status"))

ggplot(data = reservations,
       mapping = aes(x = arrival_month,
                     y = total_nights,
                     fill = booking_status)) +
  geom_bar(position = "stack", stat = "identity") +
  xlab("Date of Stay") +
  ylab("Total Nights") +
  ggtitle(label = "Reservation Cancellations With Total Night") +
  theme_bw() +
  guides(fill = guide_legend(title = "Reservation Status"))
```

```{r}

# subset data
average_cost <- mean(bookings$adr)
lower_rooms <- subset.data.frame(bookings,
      select = c("arrival_date_month", "stays_total_nights",
                 "reservation_status", "adr"))
lower_rooms <- filter(lower_rooms, lower_rooms$adr <= average_cost)

upper_rooms <- subset.data.frame(bookings,
      select = c("arrival_date_month", "stays_total_nights",
                 "reservation_status", "adr"))
upper_rooms <- filter(upper_rooms, upper_rooms$adr > average_cost)

```

Plotting data revenue for cancelled reservations and not cancelled reservations; then, comparing these findings against the reservations and total nights bar graph. 

```{r}

ggplot(data = reservations,
       aes(x = arrival_date_full, y = total_rev, fill = booking_status)) +
  geom_area() +
  xlab("Arrival date") +
  ylab("Total Revenue")


ggplot(data = bookings,
       aes(x = reservation_status_date,
           y = stays_total_nights,
           fill = reservation_status)) + 
  geom_area() +
  xlab("Status Date") +
  ylab("Total Nights")

```



Plotting data for Bookings dataset in stacked bars. 

```{r}
ggplot(data = lower_rooms,
       mapping = aes(x = arrival_date_month,
                     y = stays_total_nights,
                     fill = reservation_status)) +
  geom_bar(position = "stack", stat = "identity") +
  xlab("Month of Stay") +
  ylab("Total Nights") +
  ggtitle(label = "Lower Price Rooms Cancellations") +
  theme_bw() +
  guides(fill = guide_legend(title="Booking Status"))

ggplot(data = upper_rooms,
       mapping = aes(x = arrival_date_month,
                     y = stays_total_nights,
                     fill = reservation_status)) +
  geom_bar(position = "stack", stat = "identity") +
  xlab("Date of Stay") + ylab("Total Nights") +
  ggtitle(label = "Upper Price Room's Cancellations") +
  theme_bw() + guides(fill = guide_legend(title="Booking Status")) 

ggplot(data = bookings,
       mapping = aes(x = previous_cancellations,
                     y = arrival_date_week_number,
                     fill = customer_type)) +
  geom_tile()
```