---
title: "Price Prediction"
output: html_notebook
---

```{r}
  source("clean_data.R")
  bookings <- get_cleaned_bookings()          # Supplementary data set
  reservations <- get_cleaned_reservations()  # Primary data set
  
  reservations$order <- rownames(reservations) # for reconstruction after subsetting
  library(ggplot2)
```

```{r}
# Fitting / Predicting split by room type. Implementation is not very r-like
# but hey, on a schedule.

fit_models <- function(formula) {
  rooms <- levels(reservations$room_type_reserved)
  models <- vector("list", length(rooms))
  names(models) <- rooms
  
  for(room in rooms) {
    data <- subset(reservations, room_type_reserved == room)
    models[[room]] <- lm(formula, data)
  }
  return(models)
}

make_predictions <- function(models, pop = reservations) {
  predicted_price <- seq_len(nrow(pop))
  predictions <- data.frame()
  for(room_type in levels(pop$room_type)){
    model <- models[[room_type]]
    data <- subset(pop, room_type_reserved == room_type)
    data$predicted_price <- predict(model, data)
    predictions <- rbind(predictions, data)
  }
  return(predictions[order(predictions$order), ])
}
```


```{r}
# Evaluation utilities for a model
plot_model <- function(fit_models, pop = reservations) {
  predictions <- make_predictions(fit_models, pop)
  # manually make the plot square
  lower = min(min(predictions$predicted_price), min(predictions$avg_price_per_room))
  upper = max(max(predictions$predicted_price), max(predictions$avg_price_per_room))
  
  ggplot(predictions[sample(nrow(predictions)), ]) +
    geom_point(mapping = aes(x = avg_price_per_room, 
                             y = predicted_price,
                             col = room_type_reserved)) +
    geom_abline(slope = 1, intercept = 1) +
    ggtitle("Predicted vs Actual prices") +
    xlab("Actual Price") +
    ylab("Predicted Price") +
    coord_fixed(xlim = c(lower, upper), ylim = c(lower, upper)) +
    theme(aspect.ratio = 1)
}

```

In this document we look to build a regression model for the average price. We 
might suppose that the type of room booked would heavily impact the average price:


```{r}
  ggplot(data = reservations, mapping = aes(x = avg_price_per_room,
                                            y = room_type_reserved,
                                            fill = room_type_reserved)) +
    geom_violin() +
    geom_boxplot(alpha = 0) +
    ggtitle("Average Price by Room Type") +
    xlab("Average Nightly Price") +
    ylab("Room Type")
```

We do see that the distribution of price by room time looks different. However,
there is still a great deal of variance within each group, so predicting solely
on room type would show poor results.

We also see a few outliers, which will be removed- we aren't trying to catch
every data point.
```{r}
  removed_outliers <- data.frame()
  for (room_type in levels(reservations$room_type_reserved)){
    room_data <- subset(reservations, room_type_reserved == room_type)
    lower = mean(room_data$avg_price_per_room) - 1.5 * IQR(room_data$avg_price_per_room)
    upper = mean(room_data$avg_price_per_room) + 1.5 * IQR(room_data$avg_price_per_room)
    room_data <- subset(room_data, avg_price_per_room < upper & avg_price_per_room > lower)
    removed_outliers <- rbind(removed_outliers, room_data)
  }
  reservations <- removed_outliers
```

With the outliers removed, this is what the dataset looks like. **Note that
for some of the populations, a number of observations with prices around 0 have
been dropped: we may be missing that case.**

```{r}
  ggplot(data = reservations, mapping = aes(x = avg_price_per_room,
                                            y = room_type_reserved,
                                            fill = room_type_reserved)) +
    geom_violin() +
    geom_boxplot(alpha = 0) +
    ggtitle("Average Price by Room Type") +
    xlab("Average Nightly Price") +
    ylab("Room Type")
```

Since the populations look so different we will be fitting a linear model to each
group independently. This is further justified when we look at the behaviors for
some of the variables like number of nights stayed:

```{r}
  ggplot(data = reservations,
        mapping = aes(x = total_nights, y = avg_price_per_room, col = room_type_reserved)) +
    geom_point() +
    geom_smooth(method = "lm", formula = y~x) +
    ggtitle("Average Price by Nights Stayed") +
    ylab("Average Price per Room") + 
    xlab("Total Nights (Weekend + Week)")
```
The relationship of total nights stayed to the average price is different: 
for groups 3 and 7, longer stays cost more, but the other groups are largely 
unnafected.

```{r}
  # janky special case for when we are fitting only on 
  models_nights = fit_models(avg_price_per_room ~ 
                                  (no_of_week_nights + no_of_weekend_nights))
  plot_model(models_nights)
  summaries <- lapply(models_nights, summary)
  print(summaries)
```


Prior knowledge of hotels might also suggest that the number of people would 
affect price, with adults and children possibly being priced separately.

```{r}
  ggplot(data = reservations, mapping = aes(x = no_of_adults, y = avg_price_per_room, col = room_type_reserved)) + 
    geom_jitter() 

  ggplot(data = reservations, mapping = aes(x = no_of_children, y = avg_price_per_room, col = room_type_reserved)) + 
    geom_jitter()
```
We see that this question is not fully independent of asking the room type- for
instance, children-only reservations use almost exclusively room type 2- but there
is enough of a shift even accounting for room type to justify adding the number
of participants to the model.

After building our linear model for these and other relevant factors,
we get:

```{r}
  # janky special case for when we are fitting only on 
  models_final = fit_models(avg_price_per_room ~ 
                                  (no_of_week_nights +
                                   no_of_weekend_nights +
                                   arrival_date_full +
                                   no_of_children +
                                   no_of_adults +
                                   market_segment_type + 
                                   type_of_meal_plan +
                                   no_of_special_requests +
                                   required_car_parking_space +
                                   lead_time
                                  ))
  plot_model(models_final)
  summaries_final <- lapply(models_final, summary)
  print(summaries_final)
```

The fit is better for some room types than others. For room type 3, for example, 
the fit is perfect:

```{r}
  plot_model(models_final, subset(reservations, room_type_reserved == "Room_Type 3"))
```

```{r}
table(reservations$room_type_reserved)
```