---
title: "Price Prediction"
output: html_notebook
---

```{r}
  source("clean_data.R")
  bookings <- get_cleaned_bookings()          # Supplementary data set
  reservations <- get_cleaned_reservations()  # Primary data set
  
  library(ggplot2)
```

```{r}
# Evaluation utilities for a model
plot_model <- function(fit_model) {
  predictions <- predict(fit_model, reservations)
  # manually make the plot square
  lower = min(min(predictions), min(reservations$avg_price_per_room))
  upper = max(max(predictions), max(reservations$avg_price_per_room))
  
  ggplot() +
    geom_point(mapping = aes(x = reservations$avg_price_per_room, 
                             y = predictions,
                             col = reservations$room_type_reserved)) +
    geom_abline(slope = 1, intercept = 1) +
    ggtitle("Predicted vs Actual prices") +
    xlab("Actual Price") +
    ylab("Predicted Price") +
    coord_fixed(xlim = c(lower, upper), ylim = c(lower, upper)) +
    theme(aspect.ratio = 1)
}

```

In this document we look to build a regression model for the average price. We 
might suppose that the type of room booked would heavily impact the average price:


```{r}
  ggplot(data = reservations, mapping = aes(x = avg_price_per_room,
                                            y = room_type_reserved,
                                            fill = room_type_reserved)) +
    geom_violin() +
    geom_boxplot(alpha = 0) +
    ggtitle("Average Price by Room Type") +
    xlab("Average Nightly Price") +
    ylab("Room Type")
```

We do see that the distribution of price by room time looks different. However,
there is still a great deal of variance within each group, so predicting solely
on room type would show poor results.

We also see a few outliers, which will be removed- we aren't trying to catch
every data point.
```{r}
  removed_outliers <- data.frame()
  for (room_type in levels(reservations$room_type_reserved)){
    room_data <- subset(reservations, room_type_reserved == room_type)
    lower = mean(room_data$avg_price_per_room) - 1.5 * IQR(room_data$avg_price_per_room)
    upper = mean(room_data$avg_price_per_room) + 1.5 * IQR(room_data$avg_price_per_room)
    room_data <- subset(room_data, avg_price_per_room < upper & avg_price_per_room > lower)
    removed_outliers <- rbind(removed_outliers, room_data)
  }
  reservations <- removed_outliers
```

With the outliers removed, this is what the dataset looks like. **Note that
for some of the populations, a number of observations with prices around 0 have
been dropped: we may be missing that case.**

```{r}
  ggplot(data = reservations, mapping = aes(x = avg_price_per_room,
                                            y = room_type_reserved,
                                            fill = room_type_reserved)) +
    geom_violin() +
    geom_boxplot(alpha = 0) +
    ggtitle("Average Price by Room Type") +
    xlab("Average Nightly Price") +
    ylab("Room Type")
```

```{r}
  room_type_model = lm(avg_price_per_room ~ room_type_reserved, reservations)
  plot_model(room_type_model)
  summary(room_type_model)
```

Prior knowledge of hotels might also suggest that the number of people would 
affect price, with adults and children possibly being priced separately.

```{r}
  ggplot(data = reservations, mapping = aes(x = no_of_adults, y = avg_price_per_room, col = room_type_reserved)) + 
    geom_jitter() 

  ggplot(data = reservations, mapping = aes(x = no_of_children, y = avg_price_per_room, col = room_type_reserved)) + 
    geom_jitter()
```
We see that this question is not fully independent of asking the room type- for
instance, children-only reservations use almost exclusively room type 2- but there
is enough of a shift even accounting for room type to justify adding the number
of participants to the model.

```{r}
  people_and_room_model <- lm(avg_price_per_room ~ (room_type_reserved + 
    no_of_adults + no_of_children), reservations)
  plot_model(people_and_room_model)
  summary(people_and_room_model)
```
Visually better. (TODO: implement metrics like MSE). 

Our data was taken over a duration, during which hotel prices may have shifted.
```{r}
  ggplot(data = reservations, mapping = aes(x = arrival_date_full,
                                            y = avg_price_per_room,
                                            col = room_type_reserved)) +
    geom_point() +
    geom_smooth(method="lm", formula = y~x)
```

Price is clearly time-dependent, although the relationship may not strictly speaking
be linear. Nevertheless, add it to the model.

```{r}
  model_with_time <- lm(avg_price_per_room ~
                         (room_type_reserved + no_of_adults + no_of_children + arrival_date_full),
                        reservations)
  plot_model(model_with_time)
  summary(model_with_time)
```

```{r}
  ggplot(data = reservations,
        mapping = aes(x = total_nights, y = avg_price_per_room, col = room_type_reserved)) +
    geom_point() +
    geom_smooth(method = "lm", formula = y~x)
```
Matters to some but not much to others.

```{r}
  model_with_duration <- lm(avg_price_per_room ~
                         (room_type_reserved +
                            no_of_adults +
                            no_of_children +
                            arrival_date_full +
                            no_of_week_nights +
                            no_of_weekend_nights +
                            market_segment_type + 
                            type_of_meal_plan + 
                            no_of_special_requests + 
                            lead_time +
                            required_car_parking_space +
                            no_of_previous_cancellations +
                            booking_status),
                        reservations)
  plot_model(model_with_duration)
  summary(model_with_duration)
```














